---
title: "Regression Project _ Motor Trend"
author: "Luoning"
date: "5/15/2020"
output:
  html_document:
    df_print: paged
---

## 1.Synopsis

The work is based on data of Motor Trend, a magazine about the automobile industry. The data set of a collection of cars was reviewed. And the relationship between a set of variables and miles per gallon (MPG) (outcome) was analyzed. Two particular questions were answered: 1) whether an automatic or manual transmission is better for MPG; 2) whether the the MPG difference between automatic and manual transmissions can be quantified.

## 2.Data download and process

```{r load, cache=TRUE}
# Datasets
library(datasets)
data(mtcars)

# View of the data
head(mtcars)

str(mtcars)
```
The data frame has 32 observatoins on 11 variables.  
[, 1]	mpg	Miles/(US) gallon  
[, 2]	cyl	Number of cylinders  
[, 3]	disp	Displacement (cu.in.)  
[, 4]	hp	Gross horsepower  
[, 5]	drat	Rear axle ratio  
[, 6]	wt	Weight (1000 lbs)  
[, 7]	qsec	1/4 mile time  
[, 8]	vs	Engine (0 = V-shaped, 1 = straight)  
[, 9]	am	Transmission (0 = automatic, 1 = manual)  
[,10]	gear	Number of forward gears  
[,11]	carb	Number of carburetors  

```{r gplot, cache=TRUE, warning=FALSE}
library(dplyr)
mtcars<-mutate(mtcars,Transmission=am)
mtcars$Transmission[mtcars$Transmission==1]<-'Manual'
mtcars$Transmission[mtcars$Transmission==0]<-'Automatic'
mtcars$Transmission<-factor(paste(mtcars$Transmission))
# Plot the mpg of manual vs automatic
boxplot(mtcars$mpg ~ mtcars$Transmission, data=mtcars, outpch = 19, ylab="MPG", xlab="Transimssion", main="MPG vs Transmission")
```
  
From visuallization, the mpg of manual transmission is higher than that of automatic transmission. Now let's evaluate the significance.
```{r ttest, cache=TRUE}
auto<-mtcars[mtcars$Transmission=='Automatic',]
manual<-mtcars[mtcars$Transmission=='Manual',]
t.test(auto$mpg,manual$mpg)
```
The t value is negative and the confidence interval is absolutely below zero, which means the hypothesis that the automatic and manual transmission are the same (t=0) is rejected. The difference in the mpg of automatic and manual transmission is significant, and the mpg of automatic is lower than manual.

## Regression Model
To quantify the difference between the automatic and manual transmission. The regression model is applied to evaluate the correlation of all other variables vs. transmission and mpg relationship.

```{r lm,cache=TRUE}
fit1<-lm(mpg~am,mtcars)
summary(fit1)
```
The p-value is low but the R value is only 0.3385, which means other variables may influence the mpg. Now let's consider other variables and perform the multivariable linear regression.
```{r mvlm2,cache=TRUE}
fit2<-lm(mpg~am+cyl,mtcars)
anova(fit1,fit2)
```
The p-value is low and the cyl is significant, which means the addition of cyl as a variable is necessary. We can add more variables to check the significance.
```{r mvlm2.1,cache=TRUE}
fit2.1<-lm(mpg~am+cyl+disp+hp+drat,mtcars)
anova(fit2,fit2.1)
```
The addition of displacement (disp), Gross horsepower (hp) and Rear axle ratio (drat) is not necessary.
```{r mvlm2.2,cache=TRUE}
fit2.2<-lm(mpg~am+cyl+wt+qsec+vs,mtcars)
anova(fit2,fit2.2)
```
```{r mvlm2.3,cache=TRUE}
fit2.3<-lm(mpg~am+cyl+gear+carb,mtcars)
anova(fit2,fit2.3)
```
The new variables  weight (wt), 1/4 mile time (qsec) and Engine (vs) has at least one variable necessary. And at least one in number of forward gears (gear) and number of carburetors (carb) is necessary.
```{r mvlm2.4,cache=TRUE}
fit2.4<-lm(mpg~am+cyl+wt,mtcars)
fit2.5<-lm(mpg~am+cyl+wt+qsec,mtcars)
fit2.6<-lm(mpg~am+cyl+wt+qsec+vs,mtcars)
anova(fit2,fit2.4,fit2.5,fit2.6)
```
Only weight is significant.
carburetors (carb) is necessary.
```{r mvlm3,cache=TRUE}
fit3<-lm(mpg~am+cyl+wt,mtcars)
fit3.1<-lm(mpg~am+cyl+wt+gear,mtcars)
fit3.2<-lm(mpg~am+cyl+wt+gear+carb,mtcars)
anova(fit3,fit3.1,fit3.2)
```
Neither of forward gears (gear) and number of carburetors (carb) is necessary.  
Therefore, transmission (am), number of cylinders (cyl) and weight (wt) play a significant role on mpg. This can also be seen from the plots of mpg vs. number of cylinders and weight.
```{r mvlm4,cache=TRUE}
summary(fit3)
par(mfrow=c(1,2))
boxplot(mtcars$mpg ~ mtcars$cyl, data=mtcars, outpch = 19, ylab="MPG", xlab="Number of cylinders", main="MPG vs Cylinder")
plot(mtcars$mpg ~ mtcars$wt, data=mtcars, outpch = 19, ylab="MPG", xlab="Weight", main="MPG vs Weight")
```
  
## Residual and Diagnostics
Multivariable regression model residuals
```{r Res,cache=TRUE}
par(mfrow=c(2,2))
plot(fit3)
par(mfrow=c(1,1))
```
From the above plots, we can make the following observations:  
The residuals appear to be randomly scattered on the plot and verify the independence condition.  
The points in Q-Q plot mostly fall on the line which indicates the normally distributed residuals.  
The Scale-Location plot consists of points scattered in a constant band pattern, indicating constant variance.  
The outliers or leverage points are limited and acceptible.