---
title: "Regression Project _ Motor Trend"
author: "Luoning"
date: "5/15/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## 1.Synopsis

The work is based on data of Motor Trend, a magazine about the automobile industry. The data set of a collection of cars was reviewed. And the relationship between a set of variables and miles per gallon (MPG) (outcome) was analyzed. Two particular questions were answered: 1) whether an automatic or manual transmission is better for MPG; 2) whether the the MPG difference between automatic and manual transmissions can be quantified.

## 2.Data download and process

```{r load, cache=TRUE}
# Datasets
library(datasets)
data(mtcars)

# View of the data
head(mtcars)
```


```{r gplot, cache=TRUE, warning=FALSE}
library(dplyr)
mtcars<-mutate(mtcars,Transmission=am)
mtcars$Transmission[mtcars$Transmission==1]<-'Manual'
mtcars$Transmission[mtcars$Transmission==0]<-'Automatic'
mtcars$Transmission<-factor(paste(mtcars$Transmission))
# Plot the mpg of manual vs automatic
boxplot(mtcars$mpg ~ mtcars$Transmission, data=mtcars, outpch = 19, ylab="MPG", xlab="Transimssion", main="MPG vs Transmission")
```
  
From visuallization, the mpg of manual transmission is higher than that of automatic transmission. Now let's evaluate the significance.
```{r ttest, cache=TRUE}
auto<-mtcars[mtcars$Transmission=='Automatic',]
manual<-mtcars[mtcars$Transmission=='Manual',]
t.test(auto$mpg,manual$mpg)
```
The t value is negative and the confidence interval is absolutely below zero, which means the hypothesis that the automatic and manual transmission are the same (t=0) is rejected. The difference in the mpg of automatic and manual transmission is significant, and the mpg of automatic is lower than manual.

## Regression Model
To quantify the difference between the automatic and manual transmission. The regression model is applied to evaluate the correlation of all other variables vs. transmission and mpg relationship.

```{r lm,cache=TRUE}
fit1<-lm(mpg~am,mtcars)
summary(fit1)
```
The p-value is low but the R value is only 0.3385, which means other variables may influence the mpg. Now let's consider other variables and perform the multivariable linear regression.

```{r mvlm3,cache=TRUE}
fit3<-lm(mpg~am+cyl+wt,mtcars)
summary(fit3)
```

  
## Residual and Diagnostics
Multivariable regression model residuals
```{r Res,cache=TRUE}
par(mfrow=c(2,2))
plot(fit3)
par(mfrow=c(1,1))
```
From the above plots, we can make the following observations:  
The residuals appear to be randomly scattered on the plot and verify the independence condition.  
The points in Q-Q plot mostly fall on the line which indicates the normally distributed residuals.  
The Scale-Location plot consists of points scattered in a constant band pattern, indicating constant variance.  
The outliers or leverage points are limited and acceptible.